SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python

### 2019-06-07 scipy-1.0/paper.tex
Pruned end material

Over the last three years, our sparse matrix handling internals have been
rewritten and performance has been improved. Iterating over and slicing of CSC
and CSR matrices is now faster by up to 35\% (Figure~\ref{fig:sparse-iter}),
 and the coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
conversions are now faster (Figure~\ref{fig:sparse-conv}). Importantly,
SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing the
low-level implementations leveraged by a subset of our \texttt{sparse}
offerings.

@@

\section*{Acknowledgments}

\fixme{TODO}

%Acknowledgements should be brief, and should not include thanks to anonymous
%referees and editors, or effusive comments. Grant or contribution numbers may
%be acknowledged.

@@

\fixme{TODO}

@@

\section*{Additional information}

\fixme{TODO}

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
\caption{Assessment of \texttt{cKDTree} nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn. The calculations are based on code previously reported\cite{knn-jake}, but updated to use SciPy 1.0.0, Python 3.6, NumPy 1.13.1, and Cython 0.27.3.}
\label{fig:knn-complexity}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_iteration_bench}
\caption{Airspeed velocity benchmarks for SciPy CSC and CSR matrix row iteration over roughly three years of commits prior to the release of version 1.0. Python version 2.7, Cython version 0.27.3, NumPy 1.8.2, and the indicated non-zero element density fractions were used. The partial performance regression observed shortly after the 1.0 release is likely related to a correctness fix\cite{sparse-regress}. The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
\label{fig:sparse-iter}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_coo_dia_to_csr_csc}
\caption{Airspeed velocity benchmarks for SciPy COO / DIA to CSR / CSC matrix format conversion over roughly three years of commits prior to the release of version 1.0. Python version 2.7, Cython version 0.27.3, and NumPy 1.8.2 were used. The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
\label{fig:sparse-conv}
\end{figure}

\section{Polynomial Interpolators: Additional Details}
FITPACK provides a collection of routines for calculating
smoothing splines and automatic knot selection. One issue 
with FITPACK and its Python wrappers in SciPy, \code{UnivariateSpline}
and the \code{splrep}/\code{splev} combination, is the monolithic design:
specific algorithms for construction of interpolating and 
smoothing splines and knot selection are tightly coupled 
at the Fortran level. The library offers a very
limited degree of user control over the details 
of the spline fitting it performs. 
%
Routines for manipulating piecewise polynomials and b-splines were either
not available or lacking in ease of use and/or performance characteristics.

@@

\section*{Numerical Optimization: Additional Details}
The \texttt{scipy.optimize} subpackage provides functions for the numerical solution of several classes of root finding and optimization problems:
\begin{enumerate}
\item nonlinear root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
\item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
\item linear and nonlinear sum-of-squares problems (\texttt{leastsq}, \texttt{least\textunderscore squares}, \texttt{nnls}, \texttt{lsq\textunderscore linear}, and \texttt{curve\textunderscore fit}),
\item general, linear optimization problems (\texttt{linprog}),
\item general, nonlinear, local optimization problems of a single variable (\texttt{minimize\textunderscore scalar}),
\item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
\item general, nonlinear, global optimization problems of several variables (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
\end{enumerate}

>>>>

Over the last three years, our sparse matrix handling internals have been
rewritten and performance has been improved. Iterating over and slicing of CSC
and CSR matrices is now faster by up to 35\%, % (Figure~\ref{fig:sparse-iter}),
and the coordinate (COO) / diagonal (DIA) to CSR / CSC matrix format
conversions are now faster. % (Figure~\ref{fig:sparse-conv}).
Importantly,
SuperLU\cite{superlu_ug99} was updated to version 5.2.1, enhancing the
low-level implementations leveraged by a subset of our \texttt{sparse}
offerings.

@@

@@

M.H. edited the manuscript, created figures, compiled the author list and contributions statements, and drafted the \texttt{linprog} technical improvements section.

@@

%\section*{Additional information}
%
%\fixme{TODO}

@@

%
%\begin{figure}[H]
%\centering
%\includegraphics[width=0.8\textwidth]{supporting_info/knn_complexity_confirm}
%\caption{Assessment of \texttt{cKDTree} nearest neighbor query time complexity as a function of the number of input points (N) along with similar assessment for related algorithms in scikit-learn. The calculations are based on code previously reported\cite{knn-jake}, but updated to use SciPy 1.0.0, Python 3.6, NumPy 1.13.1, and Cython 0.27.3.}
%\label{fig:knn-complexity}
%\end{figure}
%
%\begin{figure}[H]
%\centering
%\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_iteration_bench}
%\caption{Airspeed velocity benchmarks for SciPy CSC and CSR matrix row iteration over roughly three years of commits prior to the release of version 1.0. Python version 2.7, Cython version 0.27.3, NumPy 1.8.2, and the indicated non-zero element density fractions were used. The partial performance regression observed shortly after the 1.0 release is likely related to a correctness fix\cite{sparse-regress}. The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
%\label{fig:sparse-iter}
%\end{figure}
%
%\begin{figure}[H]
%\centering
%\includegraphics[width=\textwidth]{supporting_info/asv_bench/sparse/sparse_coo_dia_to_csr_csc}
%\caption{Airspeed velocity benchmarks for SciPy COO / DIA to CSR / CSC matrix format conversion over roughly three years of commits prior to the release of version 1.0. Python version 2.7, Cython version 0.27.3, and NumPy 1.8.2 were used. The values were obtained from our regularly-updated suite of performance results available online at \url{http://pv.github.io/scipy-bench/}.}
%\label{fig:sparse-conv}
%\end{figure}

\subsection*{Polynomial Interpolators: Additional Details}
FITPACK provides a collection of routines for calculating
smoothing splines and automatic knot selection. One issue 
with FITPACK and its Python wrappers in SciPy, \code{UnivariateSpline}
and the \code{splrep}/\code{splev} combination, is the monolithic design:
specific algorithms for construction of interpolating and 
smoothing splines and knot selection are tightly coupled 
at the Fortran level. The library offers a very
limited degree of user control over the details 
of the spline fitting it performs. 
%
Routines for manipulating piecewise polynomials and b-splines were either
not available or lacking in ease of use and/or performance characteristics.

@@

\subsection*{Numerical Optimization: Additional Details}
The \texttt{scipy.optimize} subpackage provides functions for the numerical solution of several classes of root finding and optimization problems:
\begin{enumerate}
\item nonlinear root finding problems (\texttt{brentq}, \texttt{brenth}, \texttt{ridder}, \texttt{bisect}, \texttt{newton}, and \texttt{root}),
\item linear sum assignment problems (\texttt{linear\textunderscore sum\textunderscore assignment}),
\item linear and nonlinear sum-of-squares problems (\texttt{leastsq}, \texttt{least\textunderscore squares}, \texttt{nnls}, \texttt{lsq\textunderscore linear}, and \texttt{curve\textunderscore fit}),
\item general, linear optimization problems (\texttt{linprog}),
\item general, nonlinear, local optimization problems of a single variable (\texttt{minimize\textunderscore scalar}),
\item general, nonlinear, local optimization problems of several variables (\texttt{minimize}), and
\item general, nonlinear, global optimization problems of several variables (\texttt{basinhopping}, \texttt{brute}, \texttt{differential\textunderscore evolution}).
\end{enumerate}

\setlength{\tabcolsep}{3pt}
